### 2019-02-10 paper.tex
Added intro, story outline

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Large-scale peer-to-peer network for mobile platforms: Challenges and Experiences\\
%\thanks{DAIS-ITA}
}

\author{\IEEEauthorblockN{Nirmit Desai, Wendy Chong, Heather Achilles, Shahrokh Daijavad}
\IEEEauthorblockA{\textit{IBM T. J. Watson Research Center} \\
%\textit{}\\
Yorktown Heights, NY, USA \\
\{nirmit.desai, wendych, hachilles, shahrokh\}@us.ibm.com}
\and
%\IEEEauthorblockN{Wendy Chong}
%\IEEEauthorblockA{\textit{IBM T. J. Watson Research Center} \\
%\textit{name of organization (of Aff.)}\\
%Yorktown Heights, NY, USA \\
%wendych@us.ibm.com}
%\and
%\IEEEauthorblockN{Heather Achilles}
%\IEEEauthorblockA{\textit{IBM T. J. Watson Research Center} \\
%\textit{name of organization (of Af}\\
%Yorktown Heights, NY, USA \\
%hachilles@us.ibm.com}
%\and
%\IEEEauthorblockN{Shahrokh Daijavad}
%\IEEEauthorblockA{\textit{IBM T. J. Watson Research Center} \\
%\textit{Penn}\\
%Yorktown Heights, NY, USA \\
%shahrokh@us.ibm.com}
\and
\IEEEauthorblockN{Thomas La Porta}
\IEEEauthorblockA{\textit{Dept. of Computer Science and Engineering} \\
\textit{Penn State University}\\
University Park, PA, USA \\
tlp@cse.psu.edu}
%\and
%\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address}
%\and
%\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address}
}

\maketitle

\begin{abstract}
Peer-to-peer (p2p) networks and Mobile ad hoc networks (MANET) have
been widely studied. However, a real-world deployment for the masses
has remained elusive. Ever-increasing density of mobile devices,
especially in urban areas, has given rise to new applications of p2p
communication. However, the modern smartphone platforms have limited
support for such communications. Further, the issues of battery life,
range, and security remain unaddressed. A key question then is, what
kinds of applications can the modern mobile platforms support and what
challenges remain? This paper identifies a class of applications and
presents a novel p2p architecture called Mesh Network Alerts (MNA) to
support them.  We describe our experiences in deploying MNA as a
real-world peer-to-peer network to millions of users for relaying
severe weather information along with the challenges faced, and the
approaches for addressing them.
\end{abstract}

\begin{IEEEkeywords}
peer-to-peer systems, mobile ad hoc network, delay-tolerant network
\end{IEEEkeywords}

\section{Introduction}
Mobile devices with programmable platforms such as android and iOS
have steadily grown over the last decade, surpassing the 2 billion
mark \footnote{https://www.statista.com/statistics/330695/number-of-smartphone-users-worldwide/}. MANETs
have been greatly studied given their decentralized nature and
potential for new applications
\cite{loo-manet-2011,perkins-ad-hoc-2001}. Most of the prior work on
p2p networks has focused on valuable analytical and simulation-based
study of MANET behavior
\cite{zhang-topology-manet-2015,marti-misbehavior-manet-2000,mauve-pos-routing-manet-2001}. Given
the outstanding practical challenges of physical nodes, real-world
implementations have been limited and have not reached mass scale
\cite{kiess-manet-impl-2007}. However, with the growth of smartphones,
large-scale real-world implementations may become feasible. This paper
describes a real-world implementation of a p2p delay-tolerant network,
called Mesh Network Alerts (MNA), for relaying severe weather
information to millions of mobile device users as part of the Weather
Channel app\footnote{https://weather.com/apps/ibm/meshnetworkalerts}
on both android and iOS platforms.

Before describing MNA, it is critical to identify applications that
need p2p communication, given pervasive Internet connectivity. Doing
so enables us to define key characteristics of such applications and
focus on the challenges in meeting them.  This paper focuses on two
separate classes of applications: communication in
(a) disaster-affected or remote areas and (b) congested networks in
densely populated areas, e.g., sports arenas.  The following are the
key characteristics in these scenarios:

\begin{itemize}
\item No secondary communication infrastructure such as WiFi
  access points to fall back on
\item Network nodes are mobile, pattern of mobility is not
  predictable
\item New information may arrive at any time
\item Trustworthy information is scarce, misinformation and
  rumours are common place
\item  Small payloads suffice in many cases and information
  stays relevant for a few minutes
\item  Device battery is a scarce resource, power supply
  for recharging may not be available
\item  Devices are owned by citizens, deployment of
  special-purpose devices is cost prohibitive
\end{itemize}

The above needs are well-recognized in the industry with several
ambitious attempts to address them, e.g., Google Loon
project\footnote{https://loon.co} and Facebook
Aquilla\footnote{https://en.wikipedia.org/wiki/Facebook\_Aquila},
though with limited impact given. Leveraging user mobile devices as
peer nodes for a large-scale deployment has been attempted before,
e.g., the Serval project \cite{gardner-stephen-serval-2011}. Serval
mesh enables p2p communication over on-device WiFi radio, but requires
root access to the device via jailbreaking. Although significant
leassons have been learned through these attempts, a mass-scale p2p
network for such applications remains elusive.

A vast majority of the literature has focused on a traditional model
of stateful and reliable networking. Specifically, the nodes maintain
connectivity with peers, routing is optimized with techniques based on
link state or distance vectors
\cite{clausen-olsr-2003,perkins-aodv-2003} focusing on optimizing the
network utilization. Given the application characteristics above, this
paper identifies main practical challenges associated with modern
device platforms and finds novel ways to overcome them.  This leads to
a new paradigm in p2p networking that is connection-less,
delay-tolerant, and zero-routing.  With extensive experiments and
experience of deploying to almost $10$ million users, MNA represents a
way forward for large-scale p2p networks.

In summary, this paper makes the following key contributions:
\begin{itemize}
\item Identification of a class of applications for p2p and their key
  characteristics
\item A deeper investigation of the practical challenges in supporting
  the above class of applications
\item A novel p2p networking implementation using multiple radio
  channels for modern mobile platforms (Android and iOS)
\item Experimental evaluation and deployment statistics
\end{itemize}

In the following, Section~\ref{sec:challenges} describes the practical
challenges. Section~\ref{sec:architecture} outlines the architectural
details of MNA along with platform-specific implementation issues for
Android and iOS. Experimental evaluation and deployment statistics are
presented in Section~\ref{sec:eval}.  A deeper look at the literature
and contrast to MNA is summarized in Section~\ref{sec:related} with
conclusions in Section~\ref{sec:conclude}.
%
\section{Practical Challenges}
\label{sec:challenges}
%
We focus on four main challenges and describe a system implementation
for addressing them. 
%
\subsection{Background execution} 
%
Even when a user is not interacting with the weather app, or worse
yet, when the device is not being used at all, the devices must
actively discover peer devices to send and receive potentially
life-saving weather information. Although modern mobile operating
systems such as Android and iOS offer APIs to discover and advertise
information to peer devices over WiFi and Bluetooth interfaces,
peer-to-peer connections do not work when the same APIs are accessed
while the app is in the background. Prior works, widely document these
challenges and take the approach of having special access on the
devices, e.g., jail-breaking or rooting
\cite{gardner-stephen-serval-2011}. We overcame these challenges via
innovative techniques, without resorting to hacks that may violate
user security or Appstore guidelines. With our technique, exchange of
information happens without making network connections. As we describe
later, this is achieved by splitting the messages into small enough
chunks so that they can be stuffed into service advertisements
themselves and broadcast over multiple advertisements in a quick
succession.
%
\subsection{Power constraints}
%
Since devices may be offline when new weather information
arrives, MNA on each device must remain active at all times to be able
to discover new information as soon as it arrives. However, the MNA
activity must keep the device battery consumption to a minimum. This
is a challenge because advertising and discovery are power-hungry
operations over the radio channels. Our approach here is
two-pronged. Firstly, via extensive experiments on a large number of
devices, we fine-tune the algorithm governing the intervals at which
discovery and advertisements occur. In a nutshell, receiving new
information during a period causes MNA to be more aggressive in
discovery and advertisement. Similarly, lack of new information for a
period makes the device less aggressive. Secondly, we allow ``wake
up'' messages to be broadcast in the network ahead of an anticipated
severe weather event. When devices receive such messages, they
schedule themselves to remain aggressive during the specified window
of time. Outside of this window, the device can afford to have long
sleep cycles and conserve power. With these techniques, our testing
shows less than 1\% battery consumption per hour on most device
models.
%
\subsection{Testing p2p networks}
%
Given the heterogeneity of devices and operating system
distributions in the market, it is quite a challenge to test whether
MNA functions as expected on a given device. Further, running test
scenarios on a p2p network at large-scale is non-trivial given that a
large number of devices need to take specific coordinated action
followed by coordinated observations to determine whether a test
passes or fails. This is not a challenge in traditional mobile
application development as the software is confined within a single
connected device. We developed test automation tools and processes
such that multiple devices can be controlled from a single test
station and follow prescribed steps to generate, send, and receives
messages to play out a test scenario. Finally, the framework allows
automated analysis of the observations to determine the test
result. This capability was instrumental in uncovering bugs at a fast
pace to meet the aggressive timeline and avoid the cost of
acquisition. Further, the automation framework is general and can be
expressly applied to test other apps in this fashion.
%
\subsection{Trust in information}
%
As devices with mesh advertise on a continuous basis, it may
be possible for a malicious attacker to listen for such advertisements
and reverse-engineer the message formats and protocols used. Then, the
attackers may generate fake messages and advertise them, e.g., a fake
tornado alert. Such false messages cannot be distinguished from the
real ones, and MNA will end up propagating them to as many devices as
possible, ``poisoning'' the network. In general, veracity of such
information cannot be independently verified in open decentralized
distributed systems and previous work on peer-to-peer networks do not
address this challenge. In our approach, since the origin of weather
information is the Weather Channel service, digital signatures can be
attached to all weather alerts broadcast from the service. The mobile
application is distributed with the corresponding public key so that
digital signatures from the service can be verified. If a message
fails such a verification, it is discarded and not forwarded any
further.

\section{System Architecture}
\label{sec:architecture}

\section{Experimental Results}
\label{sec:eval}

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/e2e_latency}}
\caption{End-to-end latency}
\label{fig:e2e}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/duplicates}}
\caption{Duplicity due to flooding}
\label{fig:dup}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[width=\columnwidth]{figs/hops}}
\caption{Average hop count}
\label{fig:hop}
\end{figure}

\section{Related Work}
\label{sec:related}

\section{Conclusions and Future Work}
\label{sec:conclude}

\section*{Acknowledgment}
This research was sponsored by the U.S. Army Research Laboratory and
the U.K. Ministry of Defence under Agree- ment Number
W911NF-16-3-0001. The views and conclusions contained in this document
are those of the authors and should not be interpreted as representing
the official policies, either expressed or implied, of the U.S. Army
Research Laboratory, the U.S. Government, the U.K. Ministry of Defence
or the U.K. Government. The U.S. and U.K. Governments are au- thorized
to reproduce and distribute reprints for Government purposes
notwithstanding any copyright notation hereon.
\bibliographystyle{abbrv} \bibliography{smartcomp19}

\end{document}